---
title: "Motor Trend Data Analysis"
author: "Ray Qiu"
date: "October 22, 2015"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
subtitle: Regression Models Course Project
---

## Executive Summary

#### This analysis uses some data science techniques to analyze the mtcars data set, and explore the relationship between a set of variables and miles per gallon (MPG) (outcome).  The key findings are:
* Manual transmission is better than automatic transmission for MPG.
* The ratio between manual and automatic transmission for MPG is 1.806099, adjusted by cyl, disp, hp, and wt.

---
#### Load required R libraries
```{r message=FALSE}
library(ggplot2)
library(gridExtra)
```

#### Load the mtcars data and perform some basic exploratory data analysis.
```{r}
data(mtcars)
# Convert cyl, am and gear to factors
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$am <- factor(mtcars$am, labels = c("automatic", "manual"))
mtcars$gear <- as.factor(mtcars$gear)
```

#### Based on Plot #1 in Appendix, we can conclude the following: Manual transmission is better than automatic transmission for MPG.

#### Let's try to see what other variables should be included in the model
```{r}
anova(lm(mpg ~ ., data = mtcars))
```

We can pick the variables that has a p-value close to 0.05, which are: cyl, disp, hp, wt, and am. 
```{r}
fit2 <- lm(mpg ~ cyl + disp + hp + wt + am, data = mtcars)
summary(fit2)
```

Now, the "Adjusted R-squared" is 0.8344, which we believe is a good fit model for the data. 

#### Answer to Question #2: Cars with Manual transmission get better MPG than Automatic transmission, and the coefficient is 1.806099, adjusted by cyl, disp, hp, and wt.

\pagebreak

## Appendix: Plots for the model
#### Plot to show the relationships between mpg and manual/automatic transmission.
```{r}
p <- ggplot(mtcars, aes(x = am, y = mpg)) + 
    ggtitle("mpg per transmission type") +
    geom_boxplot(aes(fill = am)) +
    theme(text = element_text(size=6), axis.text.x = element_text(angle = 45, hjust = 1))
grid.arrange(p, ncol = 2, nrow = 2) # Use grid to make the plot smaller (1/4 size)
```

#### Plots for the model

```{r}
mtcars <- fortify(fit2)
plot1 <- ggplot(data = mtcars, aes(x = .fitted, y = .resid)) +
    geom_hline(yintercept = 0, colour = "firebrick3") +
    geom_point() +
    geom_smooth(se = FALSE, method = loess)
plot2 <- ggplot(data = mtcars, aes(sample = .stdresid)) +
  stat_qq() +
  geom_abline(colour = "firebrick3")
plot3 <- ggplot(data = mtcars, aes(x = .fitted, y = sqrt(abs(.stdresid)))) + 
    geom_point() +
    geom_smooth(se = FALSE, method = loess)
plot4 <- ggplot(data = mtcars, aes(.hat, .stdresid)) +
    geom_vline(size = 2, colour = "white", xintercept = 0) +
    geom_hline(size = 2, colour = "white", yintercept = 0) +
    geom_point() + 
    geom_smooth(se = FALSE, method = loess)
grid.arrange(plot1, plot2, plot3, plot4, ncol = 2)
```
