---
title: "Regression Models (Quiz 3)"
author: "Ray Qiu"
date: "October 22, 2015"
output:
    html_document:
    keep_md: yes
fig.width: 6
---

#### Question 1: Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.
```{r}
fit <- lm(mpg ~ factor(cyl) + wt, data = mtcars)
round(summary(fit)$coefficient[3,1], 3)
```

#### Question 2:Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as a possible confounding variable. Compare the effect of 8 versus 4 cylinders on mpg for the adjusted and unadjusted by weight models. Here, adjusted means including the weight variable as a term in the regression model and unadjusted means the model without weight included. What can be said about the effect comparing 8 and 4 cylinders after looking at models with and without weight included?
```{r}
fit <- lm(mpg ~ factor(cyl), data = mtcars)
fit2 <- lm(mpg ~ factor(cyl) + wt, data = mtcars)
summary(fit)$coefficient[3,1]
summary(fit2)$coefficient[3,1]
```

#### Question 3: Consider the mtcars data set. Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.
```{r}
library(lmtest)
fit <- lm(mpg ~ factor(cyl) + wt, data = mtcars)
fit2 <- lm(mpg ~ factor(cyl) * wt, data = mtcars)
summary(fit)
summary(fit2)
lrtest(fit, fit2)
```

#### Question 4: Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight inlcuded in the model as lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars). How is the wt coefficient interpretted?
```{r}
lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
```

* Since 1 ton == 2000 lbs, the estimated expected change in MPG per half ton increase in weight for for a specific number of cylinders (4, 6, 8).

#### Consider the following data set: x <- c(0.586, 0.166, -0.042, -0.614, 11.72), y <- c(0.549, -0.026, -0.127, -0.751, 1.344). Give the hat diagonal for the most influential point
```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
fit <- lm(y ~ x)
hatvalues(fit)
```

#### Question 6: Consider the following data set: x <- c(0.586, 0.166, -0.042, -0.614, 11.72), y <- c(0.549, -0.026, -0.127, -0.751, 1.344). Give the slope dfbeta for the point with the highest hat value.
```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
fit <- lm(y ~ x)
# 5 has the highest hat value
dfbetas(fit)[,2][5]
```

#### Question 7: Consider a regression relationship between Y and X with and without adjustment for a third variable Z. Which of the following is true about comparing the regression coefficient between Y and X with and without adjustment for Z.

* It is possible for the coefficient to reverse sign after adjustment. For example, it can be strongly significant and positive before adjustment and strongly significant and negative after adjustment.

